### 2014-03-10 reanalysis/ComBat_original/combat_paper_dataset2_reproduced.rmd
mer

Since limma was not the tool used in Johnson et al.'s workflow, we will first need to create the corresponding analysis in Limma for the whole ComBat adjusted data set in order to make a fair comparison.
```{r}
thismat = EB3[negativeprobesfilter, ]
thismat[thismat<1]=1
thismat=normalizeBetweenArrays(log2(thismat))
Type = as.factor(sampleannotation$Type)
design = model.matrix(~0 + Type)
fit = lmFit(thismat, design)
cont.matrix = makeContrasts ( contrasts="TypeR-TypeC", levels=design)  
fit2 = contrasts.fit(fit, cont.matrix)
EB3_limma_res = eBayes(fit2)
rm(thismat, Type, Block, design, fit, cont.matrix, fit2)
print(table(qvalue(EB3_limma_res$p.value[,1])$qvalue<0.05))

@@

>>>>

Since limma was not the tool used in Johnson et al.'s workflow, we will first need to create the corresponding analysis in Limma for the whole ComBat adjusted data set in order to make a fair comparison.
```{r}
thismat = EB3[negativeprobesfilter, ]
thismat[thismat<1]=1
thismat=normalizeBetweenArrays(log2(thismat))
Type = as.factor(sampleannotation$Type)
design = model.matrix(~0 + Type)
fit = lmFit(thismat, design)
cont.matrix = makeContrasts ( contrasts="TypeR-TypeC", levels=design)  
fit2 = contrasts.fit(fit, cont.matrix)
EB3_limma_res = eBayes(fit2)
rm(thismat, Type, design, fit, cont.matrix, fit2)
print(table(qvalue(EB3_limma_res$p.value[,1])$qvalue<0.05))

@@

```{r}
source("../../commonscripts/helper_functions.r")
nshuffleddatasets = 10
datamatrices_permuted = list()
for(i in 1:nshuffleddatasets)
{
  x = shufflesamplesinbatch( rownames(sampleannotation), sampleannotation$Batch, sampleannotation$Type, c("C", "R") )
  print(  sum(sampleannotation[,"Type"] != sampleannotation[x,"Type"]  ))  # see how many changed
  datamatrices_permuted[[i]] = as.matrix(ComBat(dat=datamatrix[, x], 
                                                batch=sampleannotation$Batch, 
                                                mod=model.matrix(~as.factor(sampleannotation[colnames(datamatrix),"Type"])),  
                                                numCovs=NULL, 
                                                par.prior=TRUE, 
                                                prior.plots=FALSE))
}

permuted_pvalcounts=list()
for(i in 1:nshuffleddatasets)
{ 
  this_p = apply(datamatrices_permuted[[i]], 1 , 
                  FUN=function(x){t.test(
                                    x[sampleannotation[, "Type"]=="C"],
                                    x[sampleannotation[, "Type"]=="R"]
                                    )$p.value})

  permuted_pvalcounts[[i]] = hist(this_p, plot=FALSE, breaks=100)$counts
  print(table(p.adjust(this_p, method="fdr")<0.05))
}
```

Forslag figur brukt i artikkel.
```{r pvaluesjohnson, dev='svg', fig.width=10, fig.height=10, tidy=FALSE}
# figure used in our article
# NB! bruker orgniale data som ikke er floored eller log2!
par(mfrow=c(2, 2))

# reproduced ComBat + limma
hist(EB3_limma_res$p.value[,1], main="P-values, real data, ComBat followed by Limma", breaks=100, xlab="p-value", ylim=c(0,3000))

# batch handled in limma
hist(datamatrix_limma_res$p.value[,1], breaks=100, main="P-values, real data, batch handled by Limma", xlab="p-value" , ylim=c(0,3000))

# random ComBat + limma
hist(EBrand_limma_pvalues, main="P-values, Random data, ComBat followed by Limma", breaks=100, xlab="p-value" , ylim=c(0,3000))

# permuted labels whitin batch

### permutation sanity test
real_pvalcounts=hist(EB3_limma_res$p.value[,1], plot=FALSE, breaks=100)$counts
plot((1:20)/100, real_pvalcounts[1:20], col="red",
     main=paste("P-values, real data and ", length(permuted_pvalcounts), " permutations" ,sep=""),
      xlab="p-value", ylab="Frequency", type="l", lwd=2, 
      ylim=c(0, max(unlist( c(permuted_pvalcounts,real_pvalcounts)))))
for(i in 1:length(permuted_pvalcounts))
{
  lines((1:20)/100,permuted_pvalcounts[[i]][1:20], col="blue")
}
legend("topright", legend=c("Reproduced p-values real labels", "Permuted C and R labels"),
       text.col=c("red", "blue"))
```

### 2014-03-10 reanalysis/Fadi_Towfic_generic_medicine/towfic_et_al.rmd
mer

```{r}
datamatrices = list()
datamatrices[["real_raw"]] = datamatrix_raw#[1:1000,]

@@

 # make the axis equal.
lim =c( min(   c(min(table_s5$Fold_Change), min(foldchange_DP_N[table_s5$Probe]))  ), 
        max( c(max(table_s5$Fold_Change), max(foldchange_DP_N[table_s5$Probe])) ))

@@

```{r}
shufflesamplesinbatch = function(samplenames, batch, covariate, shufflecovariates)
{
  ret = samplenames
	b = covariate %in% shufflecovariates
	for(thisbatch in unique(batch))
	{
		a = batch== thisbatch		
		ret[a & b] = sample( samplenames[a&b],sum(a&b),  replace=FALSE)		
	}
	return(ret)
}

@@

```{r, dev='svg', fig.width=8, fig.height=8, tidy=FALSE}
real_pvalcounts=hist(limma_p_alt, plot=FALSE, breaks=100)$counts
plot((1:20)/100, real_pvalcounts[1:20], col="red",
     main="P-values, permuted real data", 
      xlab="p-value", ylab="Frequency", type="l", lwd=2, 
      ylim=c(0, max(unlist( c(permuted_pvalcounts,real_pvalcounts)))))
for(i in 1:length(permuted_pvalcounts))
{
  lines((1:20)/100,permuted_pvalcounts[[i]][1:20], col="blue")
}
legend("topright", legend=c("Reproduced p-values real labels", "Permuted DP and N labels"), text.col=c("red", "blue"))
```
### References

>>>>

```{r}
datamatrices = list()
datamatrices[["real_raw"]] = datamatrix_raw#[1:1000,] # debug

@@

 # make the axis equal.
lim =c( min(   c(min(table_s5$Fold_Change, na.rm=T), min(foldchange_DP_N[table_s5$Probe], na.rm=T))  ), 
        max( c(max(table_s5$Fold_Change, na.rm=T), max(foldchange_DP_N[table_s5$Probe], na.rm=T)) ))

@@

```{r}
source("../../commonscripts/helper_functions.r")
#shufflesamplesinbatch = function(samplenames, batch, covariate, shufflecovariates)
#{
  #ret = samplenames
	#b = covariate %in% shufflecovariates
	#for(thisbatch in unique(batch))
	#{
	#	a = batch== thisbatch		
#		ret[a & b] = sample( samplenames[a&b],sum(a&b),  replace=FALSE)		
	#}
	#return(ret)
#}

@@

```{r, dev='svg', fig.width=8, fig.height=8, tidy=FALSE}
real_pvalcounts=hist(limma_p_alt, plot=FALSE, breaks=100)$counts
plot((1:20)/100, real_pvalcounts[1:20], col="red",
     main="P-values, permuted real data", 
      xlab="p-value", ylab="Frequency", type="l", lwd=2, 
      ylim=c(0, max(unlist( c(permuted_pvalcounts,real_pvalcounts)))))
for(i in 1:length(permuted_pvalcounts))
{
  lines((1:20)/100,permuted_pvalcounts[[i]][1:20], col="blue")
}
legend("topright", legend=c("Reproduced p-values real labels", "Permuted DP and N labels"), text.col=c("red", "blue"))
```

Forslag figur brukt i artikkel.
```{r pvaluestowfic, dev='svg', fig.width=10, fig.height=10, tidy=FALSE}
# figure used in our article
# NB! bruker orgniale data som ikke er floored eller log2!
par(mfrow=c(2, 2))

thisylim = c(0, max(unlist( c(permuted_pvalcounts,real_pvalcounts))))
# reproduced ComBat + limma
hist(limma_p_alt, main="P-values, real data, ComBat followed by Limma", breaks=100, xlab="p-value", ylim=thisylim)

# batch handled in limma
hist(limma_p_woc, breaks=100, main="P-values, real data, batch handled by Limma", xlab="p-value" , ylim=thisylim)

# random ComBat + limma
hist(limma_p_rand, main="P-values, Random data, ComBat followed by Limma", breaks=100, xlab="p-value" , ylim=thisylim)

# permuted labels whitin batch

### permutation sanity test
real_pvalcounts=hist(limma_p_alt, plot=FALSE, breaks=100)$counts
plot((1:20)/100, real_pvalcounts[1:20], col="red",
     main=paste("P-values, real data and ", length(permuted_pvalcounts), " permutations" ,sep=""),
      xlab="p-value", ylab="Frequency", type="l", lwd=2, 
      ylim=thisylim)
for(i in 1:length(permuted_pvalcounts))
{
  lines((1:20)/100,permuted_pvalcounts[[i]][1:20], col="blue")
}
legend("topright", legend=c("Reproduced p-values real labels", "Permuted DP and N labels"),
       text.col=c("red", "blue"))
```

### References
